Staged Abstract Interpreters

### 2019-03-27 paper/60_cases.tex
flag
evaluation & paper

In Section~\ref{sai}, we have shown that staging an abstract interpreter is
feasible and can be systematic after correctly identifying the binding-times,
despite the fact that the abstract interpreter is intended to be imprecise and
easy to implement. In this section, we conduct several case studies to show that
this approach is also useful and widely applicable to different analyses.

\subsection{Abstract Compilation a la Staging} \label{cs_ac}

\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC)
as a new implementation technique for abstract interpretation based static
analysis \cite{Boucher:1996:ACN:647473.727587}. The idea is inspired by partial
evaluation similar to the present paper -- the program can be known statically,
the overhead of interpretation can be eliminated. In AC, the compiled analysis
can be represented by either text or closures (higher-order functions); though
the closures can be executed immediately, the textual programs need to be
compiled and loaded first.

Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a
monovariant control-flow analysis \cite{Shivers:1991:SSC:115865.115884,
Shivers:1988:CFA:53990.54007} for continuation-passing style (CPS) programs.
Since the analyzed program is written in CPS, the analyzer is essentially a
big-step control-environment abstract interpreter. Closure generation compiles
the analysis as a closure taking an environment argument. The overhead of
traversing the abstract syntax tree of input program also has been eliminated.

In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s
abstract compilation can be understood and implemented as an instance of staging
abstract interpreters. We first revisit the original implementation of abstract
compilation of 0-CFA, and then reproduce their result by simply adding stage
annotations. The generated program of our approach improves approximately the
same extent of speed, but without changing a single line of the analyzer program
(with the use of LMS). However, closure generation requires more engineering
effort, specifically a whole-program conversion on the analyzer. Moreover, as
shown in Section~\ref{staged_ds}, our approach is able to not only remove the
interpretive overhead, but also specialize the data structures used in the
analysis, for example, the environment that maps variables to sets of lambda.

@@

\subsubsection{Closure Generation}

@@

\subsubsection{Staged 0-CFA}

>>>>

In Section~\ref{sai}, we use a toy language to show that staging an abstract
interpreter is feasible and can be systematically derived from staged concrete
interpreters and unstaged abstract interpreter.
In this section, we conduct several case studies to further show that this
methodology is practically useful and widely applicable to diverse analyses.

\subsection{Abstract Compilation \`a la Staging} \label{cs_ac}

We first revisit a similar technique called \textit{abstract compilation} (AC).
AC is introduced by \citet{Boucher:1996:ACN:647473.727587} as a new
implementation technique for abstract interpretation based static analysis.
The idea is inspired by partial evaluation similar to the present paper --
the program is known statically, and the interpretive overhead can be
eliminated. In AC, the compiled analysis can be represented by either text or
closures (higher-order functions). The minute difference is that the closures
can be executed immediately, while the textual programs need to be compiled and
loaded first.

Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} showed how to compile a
monovariant control-flow analysis \cite{Shivers:1991:SSC:115865.115884,
Shivers:1988:CFA:53990.54007} for continuation-passing style (CPS) programs.
Since the analyzed program is in CPS form, the analyzer is a big-step
control-environment abstract interpreter. The abstract compilation transforms
the analysis as a closure just taking an environment argument. Therefore the
overhead of traversing the AST of input program has been eliminated.

In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s AC
(closure generation as example) can be understood and implemented as an instance
of staging abstract interpreters. We first revisit the original implementation
of abstract compilation of 0-CFA, and then reproduce their result by simply
adding stage annotations.
The generated program of our approach improves approximately the same extent of
speed, but without changing a single line of the analyzer program using
type-based staging annotations. However, AC requires more engineering effort,
i.e., rewriting the whole analyzer into the closure generation form. Moreover,
as shown in Section~\ref{staged_ds}, our approach is able to not only remove the
interpretive overhead, but also enabling several optimizations such specialized
data structures and equivalent rewritings.

@@

\paragraph{Original AC} \todo{WIP}

@@

\paragraph{AC through Staging}
